---
title: "Game Impact"
author: "Philip Bulsink"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(HockeyModel)
```

Sometimes people want to know what the results of a game will mean for their team's playoff chances. 

This is particularly important near the end of the season (~10 games left) for teams on the bubble. 

One way to predict this would be to run simulations at each of the outcomes and compare. 

For example, let's work a game from 2018-11-23:
```{r make_score_schedule}
scores<-HockeyModel::scores
scores<-scores[scores$Date > as.Date("2008-08-01"),]
scores<-scores[scores$Date < as.Date("2018-11-22"),]
max_score_date<-max(HockeyModel::scores$Date)
schedule<-rbind(HockeyModel::scores[HockeyModel::scores$Date > as.Date("2018-11-22"), c('Date', 'HomeTeam', 'AwayTeam')], 
                HockeyModel::schedule[HockeyModel::schedule$Date > max_score_date, ])
```

Now, the first game after this point in the season is: `r knitr::kable(head(schedule,1), row.names = FALSE)`.

For this point in the season, we can get our rho and m parameters to know we're simulating for the team strength as it was then:
```{r get_rho_m, message = FALSE, error = FALSE, warning=FALSE}
m<-getM(scores = scores, currentDate = as.Date("2018-11-22"))
rho<-getRho(scores = scores, m = m)
```

With these factors, we can predict the season. in particular, we'll look for the playoff odds of both Anaheim and Vancouver.
```{r season_predict, include=FALSE}
predictions<-remainderSeasonDC(nsims = 1e3, schedule = schedule, rho=rho, m=m, regress = TRUE)
ana<-as.numeric(predictions$summary_results[predictions$summary_results$Team == 'Anaheim Ducks', 'Playoffs'])
van<-as.numeric(predictions$summary_results[predictions$summary_results$Team == 'Vancouver Canucks', 'Playoffs'])
```
```{r season_predict_fake, eval=FALSE}
predictions<-remainderSeasonDC(schedule = schedule, scores = scores, rho=rho, m=m, regress = TRUE)
```

At this date, Anaheim has a playoff odds of `r round(ana*100, 2)`%, and Vancouver has a `r round(van*100, 2)`% shot at making it into the post-season.

Having predicted the season from this point, what happens if we modify things a bit?

Let's say Anaheim wins. The exact score won't matter, but we can add in a 3-2 score for a placeholder.
```{r anaheim_win}
result<-data.frame(Date = as.Date("2018-11-23"), 
                   AwayTeam = "Vancouver Canucks", 
                   AwayGoals = 2,
                   HomeTeam = "Anaheim Ducks", 
                   HomeGoals = 3,
                   OTStatus = "",
                   League = "NHL",
                   Tie = FALSE,
                   Winner = "Anaheim Ducks",
                   Loser = "Vancouver Canucks",
                   Result = 1
                   )
#add the results to scores
newscores<-rbind(scores, result)
#remove the game from schedule
newschedule<-schedule[2:nrow(schedule),]

new_predict<-remainderSeasonDC(nsims = 1e3, schedule = newschedule, scores = newscores, rho=rho, m=m, regress = TRUE)
```

With this, we can see that Anaheim's chances increased:
```{r chances}
ana_previous_chance<-as.numeric(predictions$summary_results[predictions$summary_results$Team == "Anaheim Ducks", "Playoffs"])
ana_new_chance<-as.numeric(new_predict$summary_results[new_predict$summary_results$Team == "Anaheim Ducks", "Playoffs"])
print(paste0("Was: ", round(ana_previous_chance*100, 2), "%, now: ", round(ana_new_chance*100, 2), "%, change of ", round((ana_new_chance-ana_previous_chance)*100, 2), "%."))
```

```{r include = FALSE}
van_previous_chance<-as.numeric(predictions$summary_results[predictions$summary_results$Team == "Vancouver Canucks", "Playoffs"])
van_new_chance<-as.numeric(new_predict$summary_results[new_predict$summary_results$Team == "Vancouver Canucks", "Playoffs"])
ari_previous_chance<-as.numeric(predictions$summary_results[predictions$summary_results$Team == "Arizona Coyotes", "Playoffs"])
ari_new_chance<-as.numeric(new_predict$summary_results[new_predict$summary_results$Team == "Arizona Coyotes", "Playoffs"])
```
Of course, nothing happens in isolation. The chances for a few other teams changed too. Vancouver's odds changed by `r round((van_new_chance-van_previous_chance)*100, 2)`%, but Arizona's chanced changed too, by `r round((ari_new_chance-ari_previous_chance)*100, 2)`%.

Similarly, we can calculate the result after a OT, SO result, or loss, and see the impact of each.
```{r results, include = FALSE}
home_ot<-result
home_ot$OTStatus <- "OT"
home_ot$Tie <- TRUE
home_ot$Result <- 0.75

home_so<-home_ot
home_so$OTStatus <- "SO"
home_so$Result <- 0.6

away_win<-result
away_win$HomeGoals <-2
away_win$AwayGoals <- 3
away_win$Winner <- 'Vancouver Canucks'
away_win$Loser <- 'Anaheim Ducks'
away_win$Result <- 0

away_ot <- away_win
away_ot$OTStatus = "OT"
away_ot$Tie <- TRUE
away_ot$Result <- 0.25

away_so<-away_ot
away_so$OTStatus <-"SO"
away_so$Result <- 0.4

home_ot_score<-rbind(scores, home_ot)
home_so_score<-rbind(scores, home_so)
away_win_score<-rbind(scores, away_win)
away_ot_score<-rbind(scores, away_ot)
away_so_score<-rbind(scores, away_so)

predict_home_ot<-remainderSeasonDC(nsims = 1e3, schedule = newschedule, scores = home_ot_score, rho=rho, m=m, regress = TRUE)
predict_home_so<-remainderSeasonDC(nsims = 1e3, schedule = newschedule, scores = home_so_score, rho=rho, m=m, regress = TRUE)
predict_away_win<-remainderSeasonDC(nsims = 1e3, schedule = newschedule, scores = away_win_score, rho=rho, m=m, regress = TRUE)
predict_away_ot<-remainderSeasonDC(nsims = 1e3, schedule = newschedule, scores = away_ot_score, rho=rho, m=m, regress = TRUE)
predict_away_so<-remainderSeasonDC(nsims = 1e3, schedule = newschedule, scores = away_so_score, rho=rho, m=m, regress = TRUE)

ana_home_ot_chance<-as.numeric(predict_home_ot$summary_results[predict_home_ot$summary_results$Team == "Anaheim Ducks", "Playoffs"])
ana_home_so_chance<-as.numeric(predict_home_so$summary_results[predict_home_so$summary_results$Team == "Anaheim Ducks", "Playoffs"])
ana_away_chance<-as.numeric(predict_away_win$summary_results[predict_away_win$summary_results$Team == "Anaheim Ducks", "Playoffs"])
ana_away_ot_chance<-as.numeric(predict_away_ot$summary_results[predict_away_ot$summary_results$Team == "Anaheim Ducks", "Playoffs"])
ana_away_so_chance<-as.numeric(predict_away_so$summary_results[predict_away_so$summary_results$Team == "Anaheim Ducks", "Playoffs"])

van_home_ot_chance<-as.numeric(predict_home_ot$summary_results[predict_home_ot$summary_results$Team == "Vancouver Canucks", "Playoffs"])
van_home_so_chance<-as.numeric(predict_home_so$summary_results[predict_home_so$summary_results$Team == "Vancouver Canucks", "Playoffs"])
van_away_chance<-as.numeric(predict_away_win$summary_results[predict_away_win$summary_results$Team == "Vancouver Canucks", "Playoffs"])
van_away_ot_chance<-as.numeric(predict_away_ot$summary_results[predict_away_ot$summary_results$Team == "Vancouver Canucks", "Playoffs"])
van_away_so_chance<-as.numeric(predict_away_so$summary_results[predict_away_so$summary_results$Team == "Vancouver Canucks", "Playoffs"]) 

impact<-data.frame(Scenario = c("Home Win", "Home OT Win", "Home SO Win", "Away SO Win", "Away OT Win", "Away Win"), 
                   "Home Team (Anaheim Ducks)" = c(ana_new_chance - ana_previous_chance,
                                                   ana_home_ot_chance - ana_previous_chance,
                                                   ana_home_so_chance - ana_previous_chance,
                                                   ana_away_so_chance - ana_previous_chance,
                                                   ana_away_ot_chance - ana_previous_chance,
                                                   ana_away_chance - ana_previous_chance
                   ),
                   "Away Team (Vancouver Canucks)" = c(van_new_chance - van_previous_chance,
                                                       van_home_ot_chance - van_previous_chance,
                                                       van_home_so_chance - van_previous_chance,
                                                       van_away_so_chance - van_previous_chance,
                                                       van_away_ot_chance - van_previous_chance,
                                                       van_away_chance - van_previous_chance)
                   )
impact$Home.Team..Anaheim.Ducks.<-round(impact$Home.Team..Anaheim.Ducks.*100, 2)
impact$Away.Team..Vancouver.Canucks.<-round(impact$Away.Team..Vancouver.Canucks.*100, 2)
```

```{r, echo = FALSE}
knitr::kable(impact)
```

With nsims set to `1e3`, these can be vary noisy odds. Sometimes even Vancouver's odds can go up when Anaheim wins.  For better simulations, try `1e4` or `1e5` to smooth out the simulation repeatability and lock in a value. 
