---
title: "2021 Updates"
output: rmarkdown::html_vignette
date: "2021-07-22"
vignette: >
  %\VignetteIndexEntry{2021 Updates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(HockeyModel)
library(ggplot2)
```

#Introduction
The code has gone through a lot of changes for the 2021-2022 season. This article highlights those changes.

#Model Enhancements
The most significant part of the improvements for 2021 are enhancements to the model itself. A known issue was the under-performance of the model in predicting tie games. Looking at the prevalence of goals for tie games it was quickly apparent that the model under-predicted 2-2 and 3-3 ties significantly. While normally the model performs well in predictions of 2 or 3 goals for a team, the teams are motivated to play to a draw and take the game to overtime, as it guarantees them at least one point. By adding two factors (named `theta` and `gamma`), the chances of having a 'regular time' tie game are adjusted to match the actual. 

Let's demonstrate. I'll select for games that are only included in the most recent model parameters. Then for all games that were ties (i.e. went to OT or SO for resolution) pulling the goals at which they were tied at the end of regulation time.
```{r adjust_scores}
scores<-scores %>%
    dplyr::filter(.data$GameID %in% unique(m$data$GameID)) %>%
    dplyr::mutate('Goals' = dplyr::case_when(.data$Result == 0.25 ~ .data$HomeGoals,
                                             .data$Result == 0.4  ~ .data$HomeGoals,
                                             .data$Result == 0.5  ~ .data$HomeGoals,
                                             .data$Result == 0.6  ~ .data$AwayGoals,
                                             .data$Result == 0.75 ~ .data$AwayGoals,
                                             TRUE ~ NA_integer_))
```

From then, we can calculate the ratio of goals in regular games compared to goals in tied games. The `*2` adjusts because each tie game has two teams at that goal count.
```{r goalratios}
goalratios<-sapply(0:max(scores$Goals, na.rm = TRUE), 
                   FUN = function(x) (nrow(scores[scores$Goals == x & !is.na(scores$Goals),])/nrow(scores[!is.na(scores$Goals),]))/
                     (nrow(scores[(scores$HomeGoals == x | scores$AwayGoals == x) & is.na(scores$Goals),])/nrow(scores[is.na(scores$Goals),])) ) * 2
```

If we plot the goal ratios vs. score (in blue) we can see that they're the same shape as a poisson curve!
```{r plot_goalratios}
goalratios<-data.frame("Goals" = 0:max(scores$Goals, na.rm=T), "GoalRatio" = goalratios)

ggplot(data = goalratios) + 
  geom_point(aes_string(x="Goals", y="GoalRatio"), colour="Blue") + 
  labs(x="Goals", y="Goal Ratio", title="Goal Ratios")
```

The new `theta` and `gamma` factors produce a similarly shaped curve by developing a Poisson curve with lambda value (distribution) equal to `theta` and then multiplied the distribution by `gamma` to match the goal ratios. Plotted here (in red) you can see they match fairly well.

```{r theta}
goalratios$Dist<-dpois(goalratios$Goals, lambda=HockeyModel::theta) * HockeyModel::gamma

ggplot(data = goalratios) + 
  geom_point(aes_string(x="Goals", y="GoalRatio"), colour="Blue") + 
  geom_point(aes_string(x="Goals", y="Dist"), colour = "Red")+
  labs(x="Goals", y="Goal Ratio", title="Model and Goal Ratios")
```

This could likely be better, but for now it's better than before.

#Data Sources
The model originally used a scraper that I had written for Hockey-Reference.com. While this worked well, changes in their website had, periodically, put the data unavailable until fixes were installed. The NHL has an API for accessing data directly from them, but until now I haven't put that to use. 

A package called `nhlapi` (available on [cran](https://cran.r-project.org/package=nhlapi) and [GitHub](https://github.com/jozefhajnala/nhlapi)) has helped with this extensively. Unlike some other api access packages, this one is very lightweight. The code includes a complete adoption of this package and the removal of [HockeyScrapR](https://github.com/pbulsink/HockeyScrapR) as a dependency. 

This also provides some of the previously manual data, such as series, limiting errors and having to scramble to get predictions up before games start.

#Playoff/Cup Chances
The way that playoff and Stanley Cup chances used to be calculated was based on the playoff format that the past few years had used. But both 2019-2020 and 2020-2021 used different formats (re-seeding, divisions/conferences, etc.). So the code now uses a slower, but more accurate and much more flexible. The new method is a Monte Carlo simulation, it simulates the playoffs millions of times to estimate the odds of each result.

#General Code Improvements
The past two NHL seasons have been atypical, to say the least. It has highlighted the may assumptions that were baked into the code, such as the number of games in a season, the way that playoffs are seeded, etc. Many of those are fixed in this update.

